Synergistic retrievals of ice in high clouds from elastic backscatter lidar, Ku-band radar and submillimeter wave radiometer observations

### 2023-05-22 RST/Submission/RSTManuscript.tex
Revised Manuscript

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig02.rev.png}\\
    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the height relative to the freezing level.}\label{f2}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig01.png}\\
    \caption{Mean CS IWC profiles for 18 classes derived using the k-Means clustering algorithm. For a 
    compressed but intelligible representation, three classes are shown in different colors in each panel.
    The associated 
    mean profiles derived from CS reflectivity observations using the SSRGA scattering calculations 
    and $N_w$ parameterization developed in this study are shown using symbol *. The vertical coordinate 
    is defined relative to the freezing level.}\label{f1}
\end{figure}

@@

Moreover, the SSRGA theory was developed for millimeter and 
submillimeter-wave calculations and may not be applicable at lidar's wavelength. 
Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 
package for a backscatter lidar, given that such a lidar is expected to be flown with a 
Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory 
(ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.
While uncertainties in the lidar forward model are rather complex and difficult to quantify,
uncertainties in the lidar observations may be accounted for to some extent by including a
multiplicative random noise factor in the calculated lidar backscatter values.  Specifically, the calculated
lidar backscatter values are multiplied with a log-normally distributed random variable with 0.0 mean and a
standard deviation of 0.1.  This results in uncertainties of about 10\% in the lidar backscatter values, which is
consistent the performance of  Cloud-Aerosol Transport System (CATS) lidar system reported in \cite{pauly2019}.  
Nevertheless, the lidar simulations are rather idealized in that they do not include random errors due to daytime 
solar background noise or systematic errors such as calibration accuracy or lidar ratio
assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
considered here. There are three main reasons this approach was taken: (1) it is important to
understand the limitations of combining these three datasets before factoring in 
individual sensor limitations, (2) while more accurate calculations based on more 
realistic ice particle shapes exist, they are rather incomplete and not readily available,
and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter 
calculations based on coincident PSD observations and the Mie solution and found good 
agreement, which suggests that electromagnetic properties derived from Mie calculations 
are adequate for practical applications. 
The lidar molecular backscatter and extinction
are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; 
Bodas-Salcedo et al. (2011)). COSP simulates the lidar total attenuated backscatter 
signal and scattering ratios at 532 nm for scenes with and without clouds. COSP also 
assumes cloud particles are spherical so that the backscattering phase function is 
estimated based on the effective radius (Mie theory). Despite this assumption, the 
COSP simulations agree well with CALIOP observations. To account for multiple-scattering 
in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of 
\cite{hogan2008lidar_ms}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig03.rev.png}\\
    \caption{Simulated distributions of Ku-band radar reflectivity (left) and lidar backscatter (right)
    as function of height above 
    the freezing level. Vertical lines indicated the assumed instrument sensitivities (8dBZ for the 
    radar and 5e-7 m$^{-1}$ sr$^{-1}$ for the lidar, respectively.) }\label{f3}
\end{figure}

@@

    Instrument & Frequency/Wavelength & Noise Assumptions& Sensitivity \\
    \hline
    Lidar & 524 nm & LogNormal(0,0.1) & 5e-7 m$^{-1}$sr$^{-1}$ \\
    \hline
    Radar & 13.8 GHz & 0.5dB & 8 dBZ \\
    \hline
    Radiometer & 89 GHz & 1.0 K & N/A \\
    & 183.31+/-0.2 GHz & 1.0 K & N/A \\
    & 183.31+/-1.1 GHz & 1.0 K & N/A \\
    & 183.31+/-2.8 GHz & 1.0 K & N/A \\
    & 183.31+/-4.2 GHz & 1.0 K & N/A \\
    & 183.31+/-6.8 GHz & 1.0 K & N/A \\
    & 183.31+/-11 GHz & 1.0 K & N/A \\
    & 325.15+/-1.5 GHz & 1.0 K & N/A \\
    & 325.15+/-3.5 GHz & 1.0 K & N/A \\
    & 183.31+/-9.5 GHz & 1.0 K & N/A \\
   \hline \\

@@

\section{Results}
\subsection{Radiometer-only retrievals}
As previously mentioned, submillimeter-wave radiometers are likely to provide by themselves more complete 
information about the total amount of ice in their observing volumes than lidars or Ku-band radars with 
limited sensitivity. However, radiometers observations are an integrated measure of radiative processes in 
clouds that provide little information about the vertical distribution of ice. From this perspective, an 
evaluation in terms of the ice water path (IWP) defined as the vertical integral of the IWC, i.e. 
$IWP=\int_0^{Z_{top}}IWC(z)dz$ is insightful.  Shown in Figure \ref{f4} is the distribution of IWP 
estimated from radiometer-only observations as a function of its true value. As apparent in the figure, there is good 
correlation between the retrieved and the true IWP values. The numerical value of the correlation coefficient 
is 0.92, and there is no-overall bias. That is, the mean values of retrieved IWP and true IWP values are 
equal. However, conditional biases are apparent, with overestimation of IWP for values smaller than 
100 g/$m^2$ and some underestimation for values larger than 1000 g/$m^2$.  The biases at the low end of the 
IWP range are not surprising, given that the impact caused by ice scattering on the total radiometric signal 
is small for low values of IWP and hard to distinguish from other sources of variability in radiometer 
observations. Saturation effects are most likely responsible for underestimation at the high end.  It should 
be noted that in this evaluation, only atmospheric profiles that exhibit ice detectable by the CS radar or CALIOP lidar are used. 
Therefore,
a radiometer-only estimation procedure derived from this training dataset is likely to result in significant 
overestimation if not used in conjunction with a discrimination procedure. However, such procedure is not critical 
in this study, as, in a synergistic application, the lidar observations may be used to discriminate between clear skies and ice clouds.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig04.rev.png}\\
    \caption{Frequency plot of estimated IWP derived radiometer-observations as a function of the 
    true IWP used in observations synthesis.}\label{f4}
\end{figure}
However, although the radiometer-only estimation procedure is able to estimate the integrated amount of ice in clouds fairly
well, its ability to characterize the vertical distribution of ice in clouds is limited.  Figure \ref{f5} shows the
conditional vertical distributions of the estimated and true IWC for the 18 classes described in Section 2a 
and shown in Figure \ref{f1}. As apparent in the figure, there are significant differences between the
estimated and true IWC profiles. 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig05.rev.png}\\
    \caption{True and radiometer-only retrieved conditional mean IWC for the 18 classes described in 
    Figure \ref{f1}.}\label{f5}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig06.rev.png}\\
    \caption{Same as in Figure \ref{f4}, but for the IWC gravity center.}\label{f6}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=.481\linewidth]{Figs/fig07a.rev.png}
    \includegraphics[width=.481\linewidth]{Figs/fig07b.rev.png}
    \includegraphics[width=.481\linewidth]{Figs/fig07c.rev.png}
    \includegraphics[width=.481\linewidth]{Figs/fig07d.rev.png}
    \caption{Top Row: Density plots of IWP retrievals from lidar-only (left), and radar-only (right) observations as a function
    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f7}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig08.rev.png}\\
    \caption{Same as in Figure \ref{f4}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f8}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig09.rev.png}\\
    \caption{Same as in Figure \ref{f5}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f9}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{./Figs/fig10.rev.png}\\
    \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f10}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.481\linewidth]{Figs/fig11a.rev.png}
    \includegraphics[width=.481\linewidth]{Figs/fig11b.rev.png}
    \includegraphics[width=.481\linewidth]{Figs/fig11c.rev.png}
    \includegraphics[width=.481\linewidth]{Figs/fig11d.rev.png}
    \caption{Top Row: Density plots of IWP retrievals from lidar-radiometer (left), and radar-radiometer (right) observations as a function
    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f11}
\end{figure}

@@

Further work is necessary to assess the impact of sources of uncertainties such as potential biases in the
forward models, variability in the
PSD intercept not captured by the current parameterization, differences in the instruments' footprint sizes,
and non-uniform beam filling on the retrievals of the IWC profiles.  Other sources of uncertainties that
need be considered include the potential existence of supercooled liquid water in the clouds and uncertainties
in the electromagnetic scattering properties used in the in instruments' forward models. These uncertainties
may be best investigated and mitigated through the use of high-quality observations from field campaigns
such as The Investigation of Microphysics and Precipitation for Atlantic Coast-Threatening Snowstorms (IMPACTS)
\citep{mcmurdie2022}.

>>>>

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig02.rev.png}\\
    \caption{Multiplicative factor in the 2ICE IWC-relation as a function of the height relative to the freezing level.}\label{f2}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig01.png}\\
    \caption{Mean CS IWC profiles for 18 classes derived using the k-Means clustering algorithm. For a 
    compressed but intelligible representation, three classes are shown in different colors in each panel.
    The associated 
    mean profiles derived from CS reflectivity observations using the SSRGA scattering calculations 
    and $N_w$ parameterization developed in this study are shown using symbol *. The vertical coordinate 
    is defined relative to the freezing level.}\label{f1}
\end{figure}

@@

Moreover, the SSRGA theory was developed for millimeter and 
submillimeter-wave calculations and may not be applicable at lidar's wavelength. 
Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 
package for a backscatter lidar, given that such a lidar is expected to be flown with a 
Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory 
(ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.
While uncertainties in the lidar forward model are rather complex and difficult to quantify,
uncertainties in the lidar observations may be accounted for to some extent by including a
multiplicative random noise factor in the calculated lidar backscatter values.  
Specifically, the calculated lidar backscatter values are multiplied with a log-normally distributed random variable 
with 0.0 mean and a standard deviation of 0.1. This results in uncertainties of about 10\% in the lidar backscatter values, 
conservative compared to the expected performance of the AOS lidars at 532 nm but consistent the performance of 
Cloud-Aerosol Transport System (CATS) lidar system at 1064 nm reported in \citep{pauly2019}. While the AOS lidars 
will have expected backscatter accuracies of 2-5\%, the lidar simulations in this paper are rather idealized in that 
they do not include random errors due to daytime solar background noise or systematic errors such as calibration accuracy 
or lidar ratio assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
considered here. There are three main reasons this approach was taken: (1) it is important to
understand the limitations of combining these three datasets before factoring in 
individual sensor limitations, (2) while more accurate calculations based on more 
realistic ice particle shapes exist, they are rather incomplete and not readily available,
and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter 
calculations based on coincident PSD observations and the Mie solution and found good 
agreement, which suggests that electromagnetic properties derived from Mie calculations 
are adequate for practical applications. 
The lidar molecular backscatter and extinction
are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; 
Bodas-Salcedo et al. (2011)). COSP simulates the lidar total attenuated backscatter 
signal and scattering ratios at 532 nm for scenes with and without clouds. COSP also 
assumes cloud particles are spherical so that the backscattering phase function is 
estimated based on the effective radius (Mie theory). Despite this assumption, the 
COSP simulations agree well with CALIOP observations. To account for multiple-scattering 
in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of 
\cite{hogan2008lidar_ms}.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig03.rev.png}\\
    \caption{Simulated distributions of Ku-band radar reflectivity (left) and lidar backscatter (right)
    as function of height above 
    the freezing level. Vertical lines indicated the assumed instrument sensitivities (8dBZ for the 
    radar and 5e-7 m$^{-1}$ sr$^{-1}$ for the lidar, respectively.) }\label{f3}
\end{figure}

@@

    Instrument & Frequency/Wavelength & Noise Assumptions& Sensitivity \\
    \hline
    Lidar & 532 nm & LogNormal(0,0.1) & 5e-7 m$^{-1}$sr$^{-1}$ \\
    \hline
    Radar & 13.8 GHz & 0.5dB & 8 dBZ \\
    \hline
    Radiometer & 89 GHz & 1.0 K & N/A \\
    & 183.31+/-0.2 GHz & 1.0 K & N/A \\
    & 183.31+/-1.1 GHz & 1.0 K & N/A \\
    & 183.31+/-2.8 GHz & 1.0 K & N/A \\
    & 183.31+/-4.2 GHz & 1.0 K & N/A \\
    & 183.31+/-6.8 GHz & 1.0 K & N/A \\
    & 183.31+/-11 GHz & 1.0 K & N/A \\
    & 325.15+/-1.5 GHz & 1.0 K & N/A \\
    & 325.15+/-3.5 GHz & 1.0 K & N/A \\
    & 183.31+/-9.5 GHz & 1.0 K & N/A \\
   \hline \\

@@

\section{Results}
\subsection{Radiometer-only retrievals}
As previously mentioned, submillimeter-wave radiometers are likely to provide by themselves more complete 
information about the total amount of ice in their observing volumes than lidars or Ku-band radars with 
limited sensitivity. However, radiometers observations are an integrated measure of radiative processes in 
clouds that provide little information about the vertical distribution of ice. From this perspective, an 
evaluation in terms of the ice water path (IWP) defined as the vertical integral of the IWC, i.e. 
$IWP=\int_0^{Z_{top}}IWC(z)dz$ is insightful.  Shown in Figure \ref{f4} is the distribution of IWP 
estimated from radiometer-only observations as a function of its true value. As apparent in the figure, there is good 
correlation between the retrieved and the true IWP values. The numerical value of the correlation coefficient 
is 0.92, and there is no-overall bias. That is, the mean values of retrieved IWP and true IWP values are 
equal. However, conditional biases are apparent, with overestimation of IWP for values smaller than 
100 g/$m^2$ and some underestimation for values larger than 1000 g/$m^2$.  The biases at the low end of the 
IWP range are not surprising, given that the impact caused by ice scattering on the total radiometric signal 
is small for low values of IWP and hard to distinguish from other sources of variability in radiometer 
observations. Saturation effects are most likely responsible for underestimation at the high end.  It should 
be noted that in this evaluation, only atmospheric profiles that exhibit ice detectable by the CS radar or CALIOP lidar are used. 
Therefore,
a radiometer-only estimation procedure derived from this training dataset is likely to result in significant 
overestimation if not used in conjunction with a discrimination procedure. However, such procedure is not critical 
in this study, as, in a synergistic application, the lidar observations may be used to discriminate between clear skies and ice clouds.
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig04.rev.png}\\
    \caption{Frequency plot of estimated IWP derived radiometer-observations as a function of the 
    true IWP used in observations synthesis.}\label{f4}
\end{figure}
However, although the radiometer-only estimation procedure is able to estimate the integrated amount of ice in clouds fairly
well, its ability to characterize the vertical distribution of ice in clouds is limited.  Figure \ref{f5} shows the
conditional vertical distributions of the estimated and true IWC for the 18 classes described in Section 2a 
and shown in Figure \ref{f1}. As apparent in the figure, there are significant differences between the
estimated and true IWC profiles. 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig05.rev.png}\\
    \caption{True and radiometer-only retrieved conditional mean IWC for the 18 classes described in 
    Figure \ref{f1}.}\label{f5}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig06.rev.png}\\
    \caption{Same as in Figure \ref{f4}, but for the IWC gravity center.}\label{f6}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=.481\linewidth]{fig07a.rev.png}
    \includegraphics[width=.481\linewidth]{fig07b.rev.png}
    \includegraphics[width=.481\linewidth]{fig07c.rev.png}
    \includegraphics[width=.481\linewidth]{fig07d.rev.png}
    \caption{Top Row: Density plots of IWP retrievals from lidar-only (left), and radar-only (right) observations as a function
    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f7}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig08.rev.png}\\
    \caption{Same as in Figure \ref{f4}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f8}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig09.rev.png}\\
    \caption{Same as in Figure \ref{f5}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f9}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth,angle=0]{fig10.rev.png}\\
    \caption{Same as in Figure \ref{f6}, but with the lidar and radar observations incorporated in the
    retrievals.}\label{f10}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.481\linewidth]{fig11a.rev.png}
    \includegraphics[width=.481\linewidth]{fig11b.rev.png}
    \includegraphics[width=.481\linewidth]{fig11c.rev.png}
    \includegraphics[width=.481\linewidth]{fig11d.rev.png}
    \caption{Top Row: Density plots of IWP retrievals from lidar-radiometer (left), and radar-radiometer (right) observations as a function
    of true IWP. Bottom Row: Same as in the top row but for IWC gravity center.}\label{f11}
\end{figure}

@@

Further work is necessary to assess the impact of sources of uncertainties such as potential biases in the
forward models, variability in the
PSD intercept not captured by the current parameterization, differences in the instruments' footprint sizes,
and non-uniform beam filling on the retrievals of the IWC profiles.  Other sources of uncertainties that
need be considered include the potential existence of supercooled liquid water in the clouds and uncertainties
in the electromagnetic scattering properties used in the in instruments' forward models. These uncertainties
may be best investigated and mitigated through the use of high-quality observations from field campaigns
such as The Investigation of Microphysics and Precipitation for Atlantic Coast-Threatening Snowstorms (IMPACTS)
\citep{mcmurdie2022}. To achieve its objectives, which were driven by the need to improve the understanding of snowfall processes, 
remote sensing of snow, and the prediction of banded snow structures \citep{mcmurdie2022}, IMPACTS relied on a suite 
of active and passive instruments deployed via a satellite-simulating aircraft. These included multiple radars, one of them
operating at Ku-band, a 532 nm elastic lidar, and a sub-millimeter-wave radiometer similar to the one considered in this study.
While the objectives of IMPACTS were snowstorms, the instruments used in the campaign sampled a wide range of clouds, 
including high ice clouds.  The IMPACTS observations associated with high ice clouds may be used to derive IWC estimates that may be
directly validating using "in-situ" measurements, as the high-altitude aircraft flew in coordination with a cloud penetrating
aircraft that carried cloud and ice particle probes. These data, although not fully available yet, are expected to
provide valuable information on the accuracy of the IWC retrievals from synergistic lidar, Ku-band radar and 
sub-millimeter-wave radiometer observations and enable the refinement of the retrieval algorithm formulated in this study.
As such, they are considered a priority for future work.
